Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-16 article-models-ttinternal.tex
corrected the definition of extension algebra homomorphism

\begin{defn}
An extension algebra homomorphism from $(A,P,\epsilon_0,\epsilon_1)$ to
$(B,Q,\eta_0,\eta_1)$ in context $\Gamma$ is a pair $(f_0,f_1)$ consisting of
\begin{align*}
\jhom*{\Gamma}{A}{B}{f_0}
  \\
\jfhom*{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
\end{align*}
for which the diagrams
\begin{equation}\label{eq:extalg-hom1}
\begin{tikzcd}
\ctxext{A}{P}
  \ar{r}{\jvcomp{}{f_0}{f_1}}
  \ar{d}[swap]{\epsilon_0}
& \ctxext{B}{Q}
  \ar{d}{\eta_0}
  \\
A
  \ar{r}[swap]{f_0}
& B
\end{tikzcd}
\end{equation}
and
\begin{equation}\label{eq:extalg-hom2}
\begin{tikzcd}[column sep=large]
\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
  \ar{r}{\jvcomp{}{f_1}{\jcomp{}{\epsilon_0}{f_1}}}
  \ar{d}[swap]{\epsilon_1}
& \ctxext{Q}{\jcomp{}{\eta_0}{Q}}
  \ar{d}{\eta_1}
  \\
P
  \ar{r}[swap]{f_1}
& Q
\end{tikzcd}
\end{equation}
commute judgmentally.
\end{defn}

\begin{rmk}
I suspect that if we copy this theory of extension algebras to Martin-L\"of
type theory, with the judgmental equalities replaced by identifications, with
dependent pair types rather than those strict extensions, etcetera, then
the type of $f_1$ for which these two diagrams commute is a mere proposition.
\end{rmk}

>>>>

\begin{thm}
Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
$\jfam{\Gamma}{A}$. Then the quadruple
\begin{equation*}
(\ctxwk{A}{Q},\ctxwk{A}{R},\ctxwk{A}{\eta_0},\ctxwk{A}{\eta_1})
\end{equation*}
is an extension algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{thm}

\begin{thm}
Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then the quadruple
\begin{equation*}
(\subst{x}{Q},\subst{x}{R},\subst{x}{\eta_0},\subst{x}{\eta_1})
\end{equation*}
is an extension algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{thm}

\begin{cor}
Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{\Gamma}{B}$
and let $\jhom{\Gamma}{A}{B}{f}$. Then the quadruple
\begin{equation*}
(\jcomp{A}{f}{Q},\jcomp{A}{f}{R},\jcomp{A}{f}{\eta_0},\jcomp{A}{f}{\eta_1})
\end{equation*}
is an extension algebra in context $\ctxext{\Gamma}{A}$.
\end{cor}

\begin{defn}
An extension algebra homomorphism from $(A,P,\epsilon_0,\epsilon_1)$ to
$(B,Q,\eta_0,\eta_1)$ in context $\Gamma$ is a pair $(f_0,f_1)$ consisting of
\begin{align*}
\jhom*{\Gamma}{A}{B}{f_0}
  \\
\jfhom*{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
\end{align*}
for which the diagrams
\begin{equation}\label{eq:extalg-hom1}
\begin{tikzcd}
\ctxext{A}{P}
  \ar{r}{\jvcomp{}{f_0}{f_1}}
  \ar{d}[swap]{\epsilon_0}
& \ctxext{B}{Q}
  \ar{d}{\eta_0}
  \\
A
  \ar{r}[swap]{f_0}
& B
\end{tikzcd}
\end{equation}
and
\begin{equation}\label{eq:extalg-hom2}
\begin{tikzcd}[column sep=large]
\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
  \ar{r}{\jvcomp{}{f_1}{\jcomp{}{\epsilon_0}{f_1}}}
  \ar{d}[swap]{\epsilon_1}
& \jcomp{}{f_0}{\ctxext{Q}{\jcomp{}{\eta_0}{Q}}}
  \ar{d}{\jcomp{}{f_0}{\eta_1}}
  \\
P
  \ar{r}[swap]{f_1}
& \jcomp{}{f_0}{Q}
\end{tikzcd}
\end{equation}
commute judgmentally.
\end{defn}

\begin{rmk}
To see that the upper morphism in the diagram of \autoref{eq:extalg-hom2} has
indeed the indicated codomain provided that the diagram of \autoref{eq:extalg-hom1}
commutes judgmentally, note that we have the judgmental equalities
\begin{align*}
\jcomp{}{f_1}{\jcomp{}{f_0}{\jcomp{}{\eta_0}{Q}}}
& \jdeq 
  \jcomp{}{\jvcomp{}{f_0}{f_1}}{\jcomp{}{\eta_0}{Q}}
  \tag{by \autoref{lem:composition-threesome}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\jvcomp{}{f_0}{f_1}}{\eta_0}}{Q}
  \tag{by \autoref{lem:jcomp-jcomp}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\epsilon_0}{f_0}}{Q}
  \tag{by \autoref{eq:extalg-hom1}}
  \\
& \jdeq
  \jcomp{}{\epsilon_0}{\jcomp{}{f_0}{Q}}.
  \tag{by \autoref{lem:jcomp-jcomp}}
\end{align*}
and we indeed have the morphism $\jcomp{}{\epsilon_0}{f_1}$ from 
$\jcomp{}{\epsilon_0}{P}$ to $\jcomp{}{\epsilon_0}{\jcomp{}{f_0}{Q}}$.
\end{rmk}

\begin{rmk}
I suspect that if we copy this theory of extension algebras to Martin-L\"of
type theory, with the judgmental equalities replaced by identifications, with
dependent pair types rather than those strict extensions, etcetera, then
the type of $f_1$ for which these two diagrams commute is a mere proposition.

With this notion of morphism, a term of an extension
algebra $(A,P,\epsilon_0,\epsilon_1)$ is a pair $(x_0,x_1)$ such that
$\subst{x_1}{{x_0}{\epsilon_0}}\jdeq x_0$.
\end{rmk}
